---
title: "Assignment 7 - US Births"
author: "Andreas Reiten Skundberg"
date: "27 10 2020"
output: html_document
---

# Born In The USA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
```

Created in collaboration with Alf-Henrik Heggelund & Fredrik Sverre Nilssen

```{r}

```


The data is provided by [the CDC](https://www.cdc.gov/nchs/data_access/vitalstatsonline.htm) which gives all the births in the US in the years 2017-2019. The CDC also includes a detailed user guide to help navigate the data.

![](https://rockthedesert.typepad.com/.a/6a00d8341cb2b653ef00e55386ed268833-pi)


##### Loading and Restructuring data
```{r}
Births2017<- read.csv("Birth2017")
Births2018<- read.csv("Births2018")
Births2019<- read.csv("Births2019")

Births2017 <- Births2017 %>% mutate(Year=2017)
Births2018 <- Births2018 %>% mutate(Year=2018)
Births2019 <- Births2019 %>% mutate(Year=2019)

# Adding all the births to a combined dateset

All_Births <- bind_rows(Births2017, Births2018, Births2019)

# Changing values to numeric

All_Births$Birth.Weight <- as.numeric(All_Births$Birth.Weight)
All_Births$Birth.Month <- as.numeric(All_Births$Birth.Month)
All_Births$Birthday.of.Week <- as.numeric(All_Births$Birthday.of.Week)

```

### Proportion of Boys to Girls Born in 2017-2019


```{r}
All_Births1 <- All_Births %>%
 group_by(Year,Sex.of.Infant) %>%
  mutate(count=row_number()) %>%
  filter(count==max(count)) 



ggplot(All_Births1, aes(x = Sex.of.Infant, y = count, fill = Sex.of.Infant)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Year) + 
  ylab(expression("Number of Births")) + 
  xlab(expression("Year")) + 
  labs(title = "Proportion of Boys to Girls Each Year")

  
```

##### Dataset 

The dataset shows the total number of births by each year and gender. Shown in the count column

```{r}
All_Births1
```


### Average Birthweight In Grams By Gender

```{r}
Avg <- All_Births %>%
  group_by(Year,Sex.of.Infant) %>%
  summarise(avgWeight = mean(Birth.Weight))


# Plot
ggplot(Avg, aes(x = Year, y = avgWeight, color= Sex.of.Infant)) + 
  geom_line() + 
  ylab(expression("Weight")) +
  xlab("Year") +
  labs(title = "Average Birthweight by gender")

```

##### Dataset 

The data shows the average weight of newborns in grams by each year and gender. 
Shown in the avgWeight column. 

```{r}
Avg
```

### Proportion of Boys to Girls By Day of Week

```{r}
#3 The proportion of boys to girls by day of birth.

All_Births2 <- All_Births %>%
  group_by(Year,Birthday.of.Week,Sex.of.Infant) %>%
  mutate(count=row_number()) %>%
  filter(count==max(count)) 

# Plot
ggplot(All_Births2, aes(x = Birthday.of.Week, y = count, fill = Sex.of.Infant, color = Sex.of.Infant)) +
  geom_line(stat = "identity") + 
  ylab(expression("Number of Births")) +
  labs(title = "Proportion of Boys to Girls By Day of Week") +
  facet_wrap(~Year)
```




